Lifting CDCL to Relational Domains

### 2016-10-13 contributions.tex
abstract

\paragraph{Contributions}
In this paper, we make the following contributions.
\begin{compactenum}
\item We present a novel program analysis technique that 
lifts the CDCL algorithm to relational numerical abstract domains.   
Our proposed technique is embodied in our tool called, {\em
  Titlis}. \pscmt{It's better not to create a zoo of tool names}  
To the best of our knowledge, this is the first instantiation of CDCL 
algorithm with relational abstract domains.

>>>>

\paragraph{Contributions}
In this paper, we make the following contributions.
\begin{compactenum}
\item We present a novel program analysis technique that 
lifts the CDCL algorithm to relational numerical abstract domains.   
We have implemented our technique in 2LS verification tool.  To 
the best of our knowledge, this is the first instantiation of CDCL 
algorithm with relational abstract domains.

### 2016-10-13 domains.tex
abstract

\subsection{Properties of Domain Elements for Decision and Learning}
An important property for a clause learning SAT solver 
is that each element of the partial assignments domain 
are decomposable into precisely complementable 
elements.  This property of the domain helps to 
guide the search away from the conflicting region of the 
search space.  
\rmcmt{define precise complement}
Thus, clause learning SAT solvers requires a weak complementation property
of the abstraction.  Most numerical abstract domain, such as intervals, 
octagons, polyhedra have complementation property, that is, the domain 
elements are intersections of complementable half-spaces. 
\Omit {
The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}. 
}
Abstract Domain library, such as APRON C library~\cite{apron}, 
is specialized for the abstract interpretation framework which 
does not require complementation property.  To this end, we 
implement our own template-based polyhedra domain and interval 
domain which supports complementation operator.  
%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
  meet-irreducible. It's trivial to do that with APRON. The reason was
a different one: APRON does not support all C operators, e.g. the bitwise operators.}
For example, the complementation of the octagon in Figure~\ref{octagon} can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

>>>>

\subsection{Properties of Domain Elements for Decision and Learning}
An important property for a clause learning SAT solver 
is that each element of the partial assignments domain 
are decomposable into precisely complementable 
elements.  This property of the domain helps to 
guide the search away from the conflicting region of the 
search space.  
\rmcmt{define precise complement}
Thus, clause learning SAT solvers requires a weak complementation property
of the abstraction.  Most numerical abstract domain, such as intervals, 
octagons, polyhedra have complementation property, that is, they are
representable as intersections of complementable half-spaces. 
\Omit {
The commonly used library for numerical abstract domains  
is the APRON C library~\cite{apron}.  This library is 
used for the static analysis of the numerical variables 
of a program by abstract interpretation. APRON provides a 
C API interface to various abstract domains and libraries 
such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
{\em Linear Equalities} library.  The aim of such analysis is 
to compute invariants over numerical variables in the 
program~\cite{se2011}. 
}
Abstract Domain library, such as APRON C library~\cite{apron}, 
is specialized for the abstract interpretation framework which 
does not require complementation property.  To this end, we 
implement our own template-based polyhedra domain and interval 
domain which supports complementation operator.  
%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
\Omit {
\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
  meet-irreducible. It's trivial to do that with APRON. The reason was
a different one: APRON does not support all C operators, e.g. the bitwise
operators.} } For example, the complementation of the octagon in Figure~\ref{octagon} 
can be written as disjunction of:
\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

### 2016-10-13 paper.tex
abstract

%===============================================================================
\begin{abstract}
%===============================================================================
%
  \emph{Abstract Conflict-Driven Learning} (ACDL) is a combination of
  abstract interpretation and the well-known Conflict-Driven Clause
  Learning (CDCL) algorithm that is used in SAT solvers. ACDL has been
  successfully used with non-relational abstract domains such as
  intervals to improve the performance of a solver for floating point
  arithmetic, for instance. 
%
Whereas the
algorithms typically used in SAT solvers can be rather easily
lifted from the Boolean lattice to other non-relational domains,
lifting them to relational domains is challenging.
%because of inter-"literal" dependencies
%because of the weaker properties of "literals"
%
In this paper we present the adaptations of algorithms and heuristics
for propagation, decision making and conflict learning necessary in
order to enable the use with relational domains.
%
We have implemented these algorithms inside the 2LS verification tool
for the example of the template polyhedra domain.
%
We evaluate the performance of our analyzer in comparison with program verifiers that 
uses propositional-level solvers.  Experimental evidence on a set of 
\rmcmt{N} benchmarks \rmcmt{drawn from X} shows significant reduction 
in the number of decisions and backjumping iterations along with stronger 
deductions and learned clause aided by the richer abstract domains.
%
\end{abstract}
%===============================================================================
\input{introduction}
%===============================================================================

@@

%===============================================================================
\input{experiments}
%===============================================================================

>>>>

%===============================================================================
\begin{abstract}
%===============================================================================
%
\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a combination of
abstract interpretation and the well-known Conflict-Driven Clause Learning 
(CDCL) algorithm that is used in SAT solvers.  ACDCL alternates between a 
model search that performs overapproximate deduction with constraint 
propagation and a conflict analysis that performs under-approximate 
abduction with heuristic choice.  
%
ACDCL has been successfully used with non-relational abstract domains 
such as intervals to improve the performance of a floating-point 
decision procedure.  Whereas the propositional level algorithms used 
in SAT solvers can be rather easily lifted from the boolean lattice to 
other non-relational domains, lifting them to relational domains is 
challenging because of relational properties of deductions and clause learning.
%
In this paper, we present the non-trivial adaptations of the model search and 
conflict analysis algorithms in ACDCL in order to enable the use of relational 
abstract domains.  We have implemented these algorithms inside the 2LS 
verification tool and instantiated our analysis over a template polyhedra 
domain.  We evaluate the performance of our analyzer in comparison with 
program verifiers that uses propositional-level solvers and a static 
analysis tool.  
%
Experimental evidence on a set of benchmarks drawn from bit-vector 
regression in SV-COMP'16, bit-precise software models of hardware 
circuits auto-generated from v2c tool, and several bounds checking benchmarks 
shows significant reduction in the number of decisions and backjumping iterations 
along with stronger deductions and learned clauses aided by the richer abstract 
domains.
%
\end{abstract}
%===============================================================================
\input{introduction}
%===============================================================================

@@

%===============================================================================
%\input{experiments}
%===============================================================================
